---
title: "Foraging Conservation Evidence"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readxl)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(viridis)
library(scales)
library(kableExtra)
source("~/ownCloud/projects/ConservationEvidence/BCI Color Ramp.R")
```

### Conservation Evidence

Conservation evidence actions were downloaded from [conservationevidence.org](conservationevidence.org) which included a summary of 197 identified actions that could directly benefit bat conservation, including their overall action category, and specific threat response. Luz cross-classified these actions according to the [IUCN general and specific threat criteria](https://www.iucnredlist.org/resources/threat-classification-scheme) and Teague updated a few categories that were blank. Some actions were listed under multiple categories.   

Forty seven of these actions (24%) were identified as directly related to bat foraging behavior or foraging habitat that were associated with either **Land protection & Management** or **Education & Awareness**. 

I'm not quite sure what I've gleaned from going through these. Like all other conservation actions, there are many more untested ideas than ones with accumulated evidence. 

#### Foraging Evidence
```{r cars}
datapath <- "/Users/teagueomara/Library/CloudStorage/Box-Box/- Missions & Programs/Collaborative Projects/conservation_evidence/Bio Cons SI Shortfalls Cons Sci/data and script/"

cons_evidence <- read.csv(paste0(datapath,"ConservationEvidence.csv"))%>%
  mutate(Outcome = fct_relevel(Outcome, "Beneficial",
                               "Likely beneficial", "Limited evidence", 
                               "Trade-off ", "Unlikely beneficial", "Likely harmful", "No evidence"),
         IUCN_threat = fct_relevel(IUCN_threat, "Logging & plant harvesting",
                                   "Agriculture", "Hunting & harvesting", 
                                   "Human disturbance", "Urban development", "Energy prod. & mining", 
                                   "Climate change ", "Natural sys. modifications","White-nose Syndrome", "Invasive sp","Diseases",
                                   "Light pollution", "Water pollution", "Noise pollution", "Transportation & corridors", "Multiple threats"))

foraging_ev<- cons_evidence %>% 
  filter(Threat_Response == "Foraging habitat")

foraging_ev %>% select(-Threat_Response) %>%  
  kable() %>%
  kable_styling(font_size = 12)
```

#### Study Summaries

Eighteen actions had at least one study associated with it, 9 had at least 2 studies, and two actions had more than 10 studies associated with them. Twenty-nine of the 47 (62%) had no evidence supporting them, which is consistent with the database as a whole. The tested, positive impact actions related to foraging are all over the board. Most, however, are focused on preserving or maintaining some form of intact foraging habitat. Few studies were designed to create foraging habitat or food availability.  

```{r}
foraging_ev %>% filter(Outcome != "no evidence") %>%
  group_by(Outcome) %>% 
  summarize(numberActions = n()) %>% 
  ggplot()+
  geom_bar(aes(x = Outcome, y = numberActions, fill = Outcome), stat = "identity")+
  scale_fill_manual(values = rev(bciCatCols))+
  labs(x = "", y = "number of studies",
       title = "Outcomes of Tested Conservation Actions")+
  theme_bw()+
  theme(legend.position = "none")

```

For those actions that have any number of supporting studies, we can look at the breakdown of how they are distributed across Action Categories and ICUN Threats. Most actions are related to mitigating the effects of agriculture and other modifications to ecosystems. 
```{r}
foraging_ev %>% filter(Number_of_Studies > 0) %>%
  group_by(Action_Category, IUCN_threat) %>% 
  summarize(numberActions = n()) %>% 
  ggplot()+
  geom_bar(aes(x = IUCN_threat, y = numberActions, fill = IUCN_threat), stat = "identity")+
  facet_wrap(~Action_Category)+
  #scale_fill_manual(values = rev(bciCatCols))+
  labs(x = "", y = "number of studies")+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8),
        legend.position = "none")+
  labs(title = "IUCN Threat Categories Addressed by Conservation Studies")
  
```

And we an break these into further IUCN specific threats actions targeting non-timber crops and urbanization / habitat conversion are at the top of the list.

```{r}
foraging_ev %>% filter(Number_of_Studies > 0) %>%
  group_by(Action_Category, IUCN_threat, IUCN_specific_threat) %>% 
  summarize(numberActions = n()) %>% 
  ggplot()+
  geom_bar(aes(x = IUCN_specific_threat, y = numberActions, fill = IUCN_specific_threat), stat = "identity")+
  #facet_wrap(~IUCN_threat)+
  #scale_fill_manual(values = rev(bciCatCols))+
  labs(x = "", y = "number of studies")+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8),
        legend.position = "none")+
  labs(title = "IUCN Specific Threats Addressed by Conservation Studies")
```
### Where's the Evidence?
But are there major threat categories that haven't really been tested? The majority of identified actions don't have any evidence to either support or refute them. 

Foraging-related conservation actions targeting aspects of Climate Change, Urban Development, Light Pollution, and Transportation & Corridors need additional attention. These four are all wrapped together in a broader human-dominated landscape context, and finding appropriate foraging remedies in these areas may be very important.

```{r}
foraging_ev %>% mutate(HasEvidence = ifelse(Number_of_Studies > 0, "Has Evidence", "No Evidence")) %>% 
  group_by(Action_Category, IUCN_threat, HasEvidence) %>% 
  summarize(numberActions = n()) %>% 
  ggplot()+
  geom_bar(aes(x = IUCN_threat, y = numberActions, fill = IUCN_threat), stat = "identity")+
  facet_wrap(~HasEvidence)+
  #scale_fill_manual(values = rev(bciCatCols))+
  labs(x = "", y = "number of actions",
       title = "Evidence Gaps\nIUCN Major Threats Addressed by Potential Conservation Actions")+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8),
        legend.position = "none")
```

#### Specific Actions Wihtout Evidence
Unsurprisingly there are a number of actions targeted at land use change -- from agriculture as well as other modifications. These specific conservation actions are mostly targeted at creating or maintaining various foraging habitats and creating connectivity between roost and foraging areas. There are many impediments that will limit bat commuting paths. Lights, traffic, and large open spaces may present barriers to some species, while others will prefer to commute and foraging in these wide open spaces.

```{r}
foraging_ev %>% filter(Number_of_Studies == 0) %>%
  select(-one_of(c("Threat_Response", "Number_of_Studies", "Outcome"))) %>% 
  kable() %>%
  kable_styling(font_size = 12)
```
