---
title: "Foraging Conservation Evidence"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(viridis)
library(scales)
library(kableExtra)
source("~/ownCloud/projects/ConservationEvidence/BCI Color Ramp.R")
```

### Conservation Evidence

Conservation evidence actions were downloaded from [conservationevidence.org](conservationevidence.org) which included a summary of 213 identified actions that could directly benefit bat conservation, including their overall action category, and specific threat response. Luz cross-classified these actions according to the [IUCN general and specific threat criteria](https://www.iucnredlist.org/resources/threat-classification-scheme).  

Forty seven of these actions (22%) were identified as directly related to bat foraging behavior or foraging habitat that were associated with either **Land protection & Management** or **Education & Awareness**. Eighteen actions had at least one study associated with it, 9 had at least 2 studies, and two actions had more than 10 studies associated with them.

```{r cars}
datapath <- "/Users/teagueomara/Library/CloudStorage/Box-Box/- Missions & Programs/Collaborative Projects/conservation_evidence/Bio Cons SI Shortfalls Cons Sci/data and script/"

cons_evidence <- read.csv(paste0(datapath,"ConservationEvidence.csv"))%>%
  mutate(Outcome = fct_relevel(Outcome, "Beneficial",
                               "Likely beneficial", "Limited evidence", 
                               "Trade-off ", "Unlikely beneficial", "Likely harmful", "No evidence"),
         IUCN_threat = fct_relevel(IUCN_threat, "Logging & plant harvesting",
                                   "Agriculture", "Hunting & harvesting", 
                                   "Human disturbance", "Urban development", "Energy prod. & mining", 
                                   "Climate change ", "Natural sys. modifications","White-nose Syndrome", "Invasive sp","Diseases",
                                   "Light pollution", "Water pollution", "Noise pollution", "Transportation & corridors", "Multiple threats"))

foraging_ev<- cons_evidence %>% 
  filter(Threat_Response == "Foraging habitat")

foraging_ev %>% select(-Threat_Response) %>%  
  kable() %>%
  kable_styling()

```

#### Study Summaries

Studies with the largest, positive impact are those that 


```{r}
foraging_ev %>% filter(Outcome != "no evidence") %>%
  group_by(Outcome) %>% 
  summarize(numberActions = n()) %>% 
  ggplot()+
  geom_bar(aes(x = Outcome, y = numberActions, fill = Outcome), stat = "identity")+
  scale_fill_manual(values = rev(bciCatCols))+
  labs(x = "", y = "number of studies")+
  theme_bw()
  
  
  
```

For those actions that have any number of supporting studies, we can look at the breakdown of how they are distributed across Action Categories and ICUN Threats.
```{r}
foraging_ev %>% filter(Number_of_Studies > 0) %>%
  group_by(Action_Category, IUCN_threat) %>% 
  summarize(numberActions = n()) %>% 
  ggplot()+
  geom_bar(aes(x = IUCN_threat, y = numberActions, fill = IUCN_threat), stat = "identity")+
  facet_wrap(~Action_Category)+
  #scale_fill_manual(values = rev(bciCatCols))+
  labs(x = "", y = "number of studies")+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8),
        legend.position = "none")
  
  
  
```





```{r consNeeds, echo=TRUE}
bats <- read.csv(paste0(datapath, "chiroptera_iucn_list.csv"))%>%
  mutate(redlistCategory = fct_relevel(redlistCategory, "Data Deficient",
                                       "Least Concern", "Near Threatened", 
                                       "Vulnerable", "Endangered", "Critically Endangered"))



dd <- bats %>%
  filter(redlistCategory == "Data Deficient")

cons_land_protection <- read.csv(paste0(datapath,"LandProtection.csv"));cons_species_mgmt <- read.csv(paste0(datapath,"SpeciesMgmt.csv"));cons_education <- read.csv(paste0(datapath,"EducationAwareness.csv"));
cons_lawpolicy <- read.csv(paste0(datapath,"LawPolicy.csv"));cons_livelihood_econ <- read.csv(paste0(datapath,"LivelihoodEconomic.csv"));res_taxonomy <- read.csv(paste0(datapath,"Taxonomy.csv"));
res_popdisttrends <- read.csv(paste0(datapath,"PopDistributionTrends.csv"));res_ecology <- read.csv(paste0(datapath,"LifeHistoryEcology.csv"));res_harvest <- read.csv(paste0(datapath,"HarvestUse&Livelihoods.csv"));
res_threats <- read.csv(paste0(datapath,"Threats.csv"));res_research <- read.csv(paste0(datapath,"Research.csv"))

bat_cons_needs <- bats %>%
  left_join(cons_land_protection, by = "scientificName")%>%
  left_join(cons_species_mgmt, by="scientificName")%>%
  left_join(cons_education, by="scientificName")%>%
  left_join(cons_lawpolicy, by="scientificName")%>%
  left_join(cons_livelihood_econ, by="scientificName")%>%
  replace(is.na(.), 0)%>%
  group_by(redlistCategory)%>%
  summarise(TotSpecies= n(),
            LandProtection = sum(LandProtectionMgmt)/TotSpecies,
            SpeciesMgmt= sum(SpeciesMgmt)/TotSpecies,
            EducationAwareness = sum(EdAwareness)/TotSpecies,
            LawAndPolicy = sum(LawPolicy)/TotSpecies,
            LivelihoodEcon = sum(LivelihoodEconomic)/TotSpecies)%>%
  gather(ConservationNeed, PropSpecies, LandProtection:LivelihoodEcon, factor_key=TRUE)

bat_res_needs <- bats %>%
  left_join(res_taxonomy, by="scientificName")%>%
  left_join(res_popdisttrends, by="scientificName")%>%
  left_join(res_ecology, by="scientificName")%>%
  left_join(res_harvest, by="scientificName")%>%
  left_join(res_threats, by="scientificName")%>%
  left_join(res_research, by="scientificName")%>%
  replace(is.na(.), 0)%>%
  group_by(redlistCategory)%>%
  summarise(TotSpecies= n(),
            Taxonomy = sum(Taxonomy)/TotSpecies,
            PopTrends = sum(PopDistTrends)/TotSpecies,
            Ecology = sum(LifeHistEcology)/TotSpecies,
            HarvestUse = sum(HarvestUseLivelihoods)/TotSpecies,
            Threats = sum(Threats)/TotSpecies,
            Research = sum(Research)/TotSpecies)%>%
  gather(ResearchNeed, PropSpecies, Taxonomy:Research, factor_key=TRUE)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
